---
title: "How to run SQL in Quarto"
description: "Configure to run SQL sentences inside Quarto Documents"
date: "2025-06-08"
categories: [Quarto, SQL, R]
format: 
  html:
    toc: true
    code-fold: false
    number-sections: true
    toc-title: Contents
    toc-location: left
execute: 
  enabled: true
  freeze: auto  # re-render only when source changes
engine: knitr
knitr:
  opts_knit:
    verbose: false
---

## Install the following R packages

```{{r}}
install.packages(c('rmarkdown', 'knitr', 'DBI', 'RSQLite', 'dplyr'))
install.packages('dbplyr', dependencies = TRUE)
```

## Import the libraries in R

```{r}
#| warning: false
library(DBI)
library(RSQLite)
library(dplyr)
```

## Create the connection to the database

In this case I'm using SQLite in memory that wont create a file. And also, I'll use the dataset `mtcars` included in the R packages.

```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), 
                      dbname = ":memory:")

copy_to(dest = con, 
        df = mtcars, 
        name = "mtcars")

dbListTables(con)
```

:::{.callout-note}
In this point we are ready to start SQL executions in this Quarto document.

:::

## Select

We start applying a Select sentence on the database `mtcars` created in the step before.

```{sql}
#| connection: con
SELECT * FROM mtcars;
```

## Apply a SUM functions in one of the columns

```{sql}
#| connection: con
SELECT SUM(mpg) FROM mtcars;
```

## Apply a distinct in one column

```{sql}
#| connection: con
SELECT distinct(cyl) FROM mtcars;
```

## Personal notes

Creating this blog I discover the following things:

- `SQL` interface in quarto only accepts `SELECT` statements
- This requires `rmarkdown` and `knitr` R libraries installed
- The package `dplyr` is to use the function `copy_to()`

## References {.unnumbered}

Special thanks to the following links:

- <https://forum.posit.co/t/sql-errors-while-rendering-quarto/147671>
- <https://quarto.org/docs/computations/r.html>

